Quantifying the home ground advantage with `hyper2` and `hyper3` objects

### 2025-02-07 inst/home_advantage.Rmd
refactoring

```{r,label=optimizebaseball,cache=TRUE}
baseball_table
f <- function(l){ maxp(home_away3(baseball_table,l),give=TRUE,n=1)$likes}
o <- optimize(f,c(0.8,1.6),maximum=TRUE)
o
maxL <- f(o$maximum)
W <- maxL - f(1)    # f(1) being the likelihood of the null
W
pchisq(2*W,df=1,lower.tail=FALSE)
```

@@

```{r, label=hyper3_baseball,cache=TRUE}
jj1  <- seq(from=log(o$maximum),to=log(1.9),len=7)
jj2  <- seq(from=log(o$maximum),to=log(0.8),len=7)
l <- sort(unique(exp(c(jj1,jj2))))
lam  <- sort(c(l,o$maximum))
L <- sapply(lam,f) - maxL
```

@@

```{r scottychisq,cache=TRUE}
o
maxL <- f(o$maximum)
W <- maxL - f(1)    # f(1) being the likelihood of the null
W
pchisq(2*W,df=1,lower.tail=FALSE)
```

```{r scotmax}
(H <- home_away3(home_away_table(a,give=FALSE),o$maximum))  # o$maximum ~= 1.353
maxp(H,give=1)
```

@@

```{r sfasdf, cache=TRUE}
o$maximum
mH3 <- maxp(home_away3(home_away_table(a,give=FALSE),o$maximum),give=FALSE)
round(mH3*1000)
```

@@

```{r}
mH
mH["Rangers"]
mH["home"]
mH3
lambda <- o$maximum
h2_exp <- matrix(0,4,4)
rownames(h2_exp) <- teams
colnames(h2_exp) <- teams
h3_exp <- h2_exp
n_matches <- Re(aF) + Im(aF)
for(hometeam in teams){
    for(awayteam in teams){

@@

>>>>

```{r,label=optimizebaseball,cache=TRUE}
baseball_table
f <- function(l){ maxp(home_away3(baseball_table,l),give=TRUE,n=1)$likes}
o <- optimize(f,c(0.8,1.6),maximum=TRUE)
o
lambda_baseball <- o$maximum
maxL <- f(lambda_baseball)
W <- maxL - f(1)    # f(1) being the likelihood of the null
W
pchisq(2*W,df=1,lower.tail=FALSE)
```

@@

```{r, label=hyper3_baseball,cache=TRUE}
jj1  <- seq(from=log(lambda_baseball),to=log(1.9),len=7)
jj2  <- seq(from=log(lambda_baseball),to=log(0.8),len=7)
l <- sort(unique(exp(c(jj1,jj2))))
lam  <- sort(c(l,lambda_baseball))
L <- sapply(lam,f) - maxL
```

@@

```{r scottychisq,cache=TRUE}
o
scot_max_lambda <- o$maximum
maxL <- f(scot_max_lambda)

W <- maxL - f(1)    # f(1) being the likelihood of the null
W
pchisq(2*W,df=1,lower.tail=FALSE)
```

```{r scotmax}
(H <- home_away3(home_away_table(a,give=FALSE),scot_max_lambda))  # o$maximum ~= 1.353
maxp(H,give=1)
```

@@

```{r sfasdf, cache=TRUE}
o$maximum
mH3 <- maxp(home_away3(home_away_table(a,give=FALSE),scot_max_lambda),give=FALSE)
round(mH3*1000)
```

@@

```{r}
mH
mH["Rangers"]
mH["home"]
mH3
lambda <- scot_max_lambda
h2_exp <- matrix(0,4,4)
rownames(h2_exp) <- teams
colnames(h2_exp) <- teams
h3_exp <- h2_exp
n_matches <- Re(aF) + Im(aF)
for(hometeam in teams){
    for(awayteam in teams){

@@

```{r defineexpectation2}
expectation2 <- function(probs, results){
    teams <- names(probs[names(probs) != "home"])
    h2_exp <- matrix(0,4,4)
    rownames(h2_exp) <- teams
    colnames(h2_exp) <- teams
    n_matches <- Re(aF) + Im(aF)
    for(hometeam in teams){
        for(awayteam in teams){

            h2_exp[hometeam,awayteam] <- (
                n_matches[hometeam,awayteam] * (mH[hometeam] + mH["home"])/
                (mH[hometeam] + mH[awayteam] + mH["home"])
            )

            h2_exp[hometeam,awayteam] %<>%
                add(1i*
                    n_matches[hometeam,awayteam] * (mH[awayteam])/
                    (mH[hometeam] + mH[awayteam] + mH["home"])
                    )
        }
    }
    return(h2_exp)
}
round(expectation2(mH,aF),1)
```

```{r defineexp3}

expectation3 <- function(probs, results, lambda){
    teams <- names(probs)
    h3_exp <- matrix(0,4,4)
    rownames(h3_exp) <- teams
    colnames(h3_exp) <- teams
    n_matches <- Re(aF) + Im(aF)
    for(hometeam in teams){
        for(awayteam in teams){

            h3_exp[hometeam,awayteam] %<>%
                add(1i*
                    n_matches[hometeam,awayteam] * (mH3[awayteam])/
                    (lambda* mH3[hometeam] + mH3[awayteam])
                    )
        }
    }
    return(h3_exp)
}

round(expectation3(mH3[c("Rangers","Celtic","Falkirk","Livingston")],aF,scot_max_lambda),1)
```
